---
title: "Predict 2012 Baseball Players' Salaries"
author: "Yulan Rong"
date: "12/10/2019"
output:
  rmarkdown::github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(glmnet)
library(car)
library(leaps)

```

##### The goal of this project is to predict salaries of baseball players by analyzing data from the 2012 baseball season from Sean Lahmann's database.

#### Data Exploration and Feature Creation 
```{r}
load(url("http://www.stat.berkeley.edu/users/nolan/data/baseball2012.rda"))
head(baseball)
dim(baseball)
dat = baseball[, -(1:6)]
head(dat)

AVG = dat$H / dat$AB
dat$AVG = AVG
CAVG = dat$CH / dat$CAB
dat$CAVG = CAVG

OBP = 100 * (dat$H + dat$BB) / (dat$AB + dat$BB)
dat$OBP = OBP
COBP = 100 * (dat$CH + dat$CBB) / (dat$CAB + dat$CBB)
dat$COBP = COBP

AB.per = dat$CAB / dat$years
dat$AB.per = AB.per
H.per = dat$CH / dat$years
dat$H.per = H.per
HR.per = dat$CHR / dat$years
dat$HR.per = HR.per
R.per = dat$CR / dat$years
dat$R.per = R.per
RBI.per = dat$CRBI / dat$years
dat$RBI.per = RBI.per

dat$POS[dat$POS == '2B'| baseball$POS == 'SS' | 
          baseball$POS == 'C'|baseball$POS == 'CF'] = 1
dat$POS[dat$POS != 1] = 0
dat$POS = as.factor(dat$POS)

dat$threetofive = factor(ifelse(
  (dat$years >= 3 & dat$years <= 5), 1, 0))
dat$oversix = factor(ifelse((dat$years >= 6), 1, 0))

dat$salary = log(dat$salary)
dat$CAB = log(dat$CAB)

dat$G_batting = NULL

dat = dat[!is.na(dat$salary),]

pairs(~salary + years + CHR +CAVG + CR + COBP + CAB +CH, data = dat)
ggplot(dat, aes(x = POS, y = salary)) + geom_boxplot() + 
  labs(title = "Position and salary", x = 'Position', y = 'salary')
ggplot(dat, aes(x = CAB, y = salary)) + geom_point() + 
  labs(title = "CAB and salary", x = 'CAB', y = 'salary')
ggplot(dat, aes(x = COBP, y = salary)) + geom_point() + 
  labs(title = "career on-base percentage and salary", 
       x = 'career on-base percentage', y = 'salary')

```


#### Data Analysis 
##### 1. Fit a simple model to predict salaries.
```{r}
dat$log.CR = log(dat$CR + 1)

simpleModel = lm(salary ~ threetofive + oversix + log.CR + 
                   oversix:log.CR + threetofive:log.CR,, data = dat)
summary(simpleModel)

plot(simpleModel$fitted.values, simpleModel$residuals, xlab = "fitted values", ylab = "residuals")
abline(h = 0, col = 'red')
# we can see the variance of the residual is not constant since 
# when fitted value increases, residuals spread out.

qqnorm(simpleModel$residuals)
qqline(simpleModel$residuals)
# from the residual qq plot we can see that the points on the left side
# are not on the line. So the residuals are not enough normally
# distributed. this model is not perfect.  
```

##### 2.check for outliers, leverage and influential observation.
```{r}
# check for leverage.
leveragePlots(simpleModel)
X <- matrix(c(rep(1,421),dat$log.CR, 
              dat$threetofive, dat$oversix),421,4) 
H <- X %*% solve(t(X) %*% X) %*% t(X)
h = diag(H)
# according to Fox, A rough cutoff for noteworthy 
# hat-values is hi > 2h_hat = 2(k+1)/n.
h.mean <- (4+1)/nrow(dat)
# check if there's leverage point that exceed twice of the h. 
sortedh <- sort(h, decreasing=TRUE, index.return=TRUE) 
unusual <- sortedh$ix[sortedh$x > 2*h.mean]
ggplot(dat,aes(x = log.CR, y = salary)) +
geom_point() + geom_point(data = dat[unusual,], aes(x = log.CR, y = salary), color = "red") + labs (x = "log career run", y = "log salary", title = "Leverage in red")

# check for outliers
outlierTest(simpleModel) 
# the outlier of this model is row 500 and row 100
qqPlot(simpleModel, main = "QQ Plot")
S_E2=sum(simpleModel$residuals^2)/(nrow(dat)-4-1)
standE=simpleModel$residuals/(sqrt(S_E2)*sqrt(1-h))
# the studentized residuals for all the points.
studE=standE*sqrt((421 - 4 -2)/(421-4-1-standE^2))
abs_studE = abs(studE)
a = dat
a$abs_studE = abs_studE
# 6 observations with the largest absolute values of the studentized residuals:
head(a[order(-abs_studE),]) 
# when p =0.05, t-test= 1.649, so set +-1.649 as bounds of outliers.
sorted_studE <- sort(abs_studE, decreasing=TRUE, index.return=TRUE)
unusual_stud <- sorted_studE$ix[sorted_studE$x < -1.649 
                                | sorted_studE$x > 1.649]
ggplot(data = dat, aes(x = log.CR, y = salary),) + geom_point() + geom_point(data = dat[unusual_stud,], aes(x = log.CR, y = salary), color = "red") + labs (x = "log career run", y = "log salary", title = "outlier in red")


# H0: it's not an outlier.
Emax=max(abs_studE)
# Hypothesis test without Bonferroni correction:
p1=1-pt(Emax,421-4-2)
p1
# Hypothesis test with Bonferroni correction:
ptrue=2*421*p1
ptrue
# since ptrue is less than 0.05, so outliers exist.

# check influential observation
dis = cooks.distance(simpleModel)
temp = dat
temp$distance = dis
# 6 observations that have largest values of Cook's distance.
head(temp[order(-dis),])
sortedCook = sort(dis, decreasing = TRUE, index.return = TRUE)
criterion = 4/(421-4-1)
influntial = sortedCook$ix[sortedCook$x > criterion]
ggplot(data = dat, aes(x = log.CR, y = salary)) + geom_point() + geom_point(data = dat[influntial,], aes(x = log.CR, y = salary), color = "red") + labs (x = "log career run", y = "log salary", title = "influential in red")
```

##### 3 Fit a linear least-squares regression model
```{r}
test = lm(salary~.,data = dat)
summary(test)
# from the Pr(>|t|), I drop the variables with high Pr.
todrop = c("CR", "POS", "AVG", "SF", "X3B", "PO", "CAB", "E", "A", "SH", "OBP", "H.per")
dat1 = dat[, !(names(dat) %in% todrop)]
fullmod = lm(salary~., data = dat1)
summary(fullmod)
```

##### 4 Find 10 best models using regsubsets.
```{r}
bestfit = regsubsets(salary~., data = dat1, method = "forward", nvmax = 33, nbest = 10)
bestfitsum = summary(bestfit)
plot(bestfitsum$adjr2, xlab = "number of variables", ylab = "adjusted R squared")
plot(bestfitsum$rss, xlab = "number of variables", ylab = "RSS")
plot(bestfitsum$bic, xlab = "number of variables", ylab = "BIC")
plot(bestfit, scale = "bic")
```

##### 5 Find 5 best models using BIC
```{r}
bic = bestfitsum$bic
bic = sort(bic, decreasing = FALSE, index.return = TRUE)
best5 = head(bic$ix, 5)
# best 5 variable model for each is different
coef(bestfit, best5)
```

##### 6 Cross validation
```{r}
set.seed(200)
permutation <- sample(1:nrow(dat1))
folds <- c(rep(1:10, each = nrow(dat1)/10), 10) 
X <- model.matrix(salary ~ ., dat1)
avg_test_MSE <- rep(0,5)
for(i in 1:5){
  test_MSE <- rep(0,10)
  for(j in 1:10){
    idx_train <- permutation[folds != j]
    idx_test <- permutation[folds == j]
    vars <- bestfitsum$which[best5[i],]
    X_best_subset <- X[,vars]
    mod <- lm(dat1$salary ~ X_best_subset - 1, subset = idx_train) 
    X_test <- X_best_subset[idx_test,]
    test_predictions <- X_test %*% as.matrix(coef(mod))
    test_MSE[j] <- mean((dat1$salary[idx_test] - test_predictions)^2)
  }
  avg_test_MSE[i] <- mean(test_MSE) 
}

plot(sqrt(avg_test_MSE), xlab = "Model Ranking in BIC", ylab = "Root Mean Squared Error", pch = 20, cex = 2)
avg_test_MSE
which.min(avg_test_MSE)
# from the 10 folds cross validation, model 5 has the lowesr average 
# test MSE among the 5 best models. 
```

##### 7 LASSO model
```{r}
X = model.matrix(salary~., dat1)[, -1]
y = dat1$salary
lambda.grid = 10^seq(10, -2, length = 100)
lasso.mod = glmnet(x = X, y = y, alpha = 1)
cv.lasso.mod = cv.glmnet(x = X, y = y, alpha = 1, nfolds = 10)

plot(cv.lasso.mod)

# the best（suitable) lasso lambda, which has the min lambda:
best.lasso.lam = cv.lasso.mod$lambda.min
best.lasso.lam 
# this best lasso lambda minimizes the mean squared error. 


plot(lasso.mod, xvar = "lambda")
lines(c(log(best.lasso.lam), log(best.lasso.lam)), c(-1000, 1000), lty = "dashed", lwd = 3)

best.lasso.coefs <- predict(lasso.mod, type = 'coefficients', s = best.lasso.lam)

# the non zero coefficients:
best.lasso.coefs
# except GS, others have non zero coefficients, 
# but many of them's coefficients are close to zero.

```

##### 8 Conclusion
```{r}
min(cv.lasso.mod$cvm)
avg_test_MSE
min(avg_test_MSE)
# by comparing the CV MSE from step 6 and LASSO, LASSO has higher 
# average MSE than the best 5 models' average Test MSE, so using 
# BIC method for choosing the best model is better than using LASSO.
# Thus, the best model in BIC method, which is model 5 will have a  
# good performance to predict this dataset.


coef(bestfit, best5)[[5]] # best model among others above.
```

