---
title: "Getting Started with roller"
author: "Yulan Rong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to roller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(roller)
```

## Rolling an Object

The package "roller" implements functions for simulating rolling an object like a coin or a die multiple times, and to visualize the relative frequencies using barchart for rolls.


## Creating a Device

The first step is to create a "device" object with the function `device()`. The output displays the class of the object, and a tabular display of the sides and the associated probabilities:

```{r}
fair_coin <- device() 
fair_coin
```

By default, `device()` creates a fair coin with two sides and identical probability 0.5 with object "device". 

You can change the default by using arguments 'sides' and 'prob':
```{r}
weird_die <- device(
sides = c('i', 'ii', 'iii', 'iv'), 
prob = rep(1/4, 4))
weird_die
```

You can use function `is.device()` to check if a given object is of class "device" :
```{r}
is.device(weird_die)
```

## Rolling the device

When you have created a device, you can use the function `roll()` to roll your device with a given 'times' you like：

```{r}
set.seed(123)
fair_50rolls <- roll(fair_coin, times = 50) 
fair_50rolls
```

The `summary()` function outputs an object "summary.rolls" with a list containing a data frame called freqs with 3 columns:
```{r}
summary(fair_50rolls)
```

## Plot of Relative Frequencies

You can use the function `plot()` to create a barchart with the relative 
frequencies of a series of rolls:

```{r, fig.show='hold'}
plot(fair_50rolls)
```

## Additional Method 

You can extract, replace the value of a given roll, or add more rolls:
```{r}
fair_50rolls[1]
fair_50rolls[1] <- 1
fair_50rolls[1]
summary(fair_50rolls)
fair_60rolls <- fair_50rolls + 10
summary(fair_60rolls)
```
